---
title: "Trump Logistic Regression"
author: "Greg"
date: "2024-04-23"
output: html_document
---

```{r, include=FALSE}
library(poliscidata)
library(tidyverse)
```
Coding the Variables- 

This section takes data from a survey and turns responses into binary varibles to make them easier to work with. The DV is transformed into either Trump vote or not Trump vote, and the two IVs are education level, turned into college or no college, and whether respondents think institutions in the U.S. are broken or not.

```{r}

dataset <- read.csv("Final project dataset ct.csv")

##Lets make our dummies, first Trump Vote coded as 1 or 0
dataset$trumpvote <- as.numeric(dataset$Q19=="Donald Trump")


##now the broken institutions variables Q65, Does respondent think instituions are broken in the U.S.?
dataset$broke <- dataset$Q65
dataset <- dataset %>%
  mutate(broke = case_when(
    Q65 %in% c("Strongly agree", "Somewhat agree") ~ "agree",
    Q65 %in% c("Neither agree nor disagree", "Strongly disagree", "Somewhat disagree") ~ "disagree",
    TRUE ~ NA_character_
  ))
levels(dataset$broke)=c("agree", "disagree")
levels(dataset$broke)
#check the new variable, what prop of our population is in each new bin
describe(dataset$broke)

##Now code Q57, party ID into a binary
dataset <- dataset %>%
  mutate(partyid = case_when(
    Q57 %in% c("Democrat", "Independent, but closer to Democrat") ~ "democrat",
    Q57 %in% c("Republican", "Independent", "I'm not really a political person", "Independent, but closer to Republican") ~ "notdemocrat",
    TRUE ~ NA_character_
  ))
levels(dataset$partyid)=c("democrat", "notdemocrat")

describe(dataset$partyid)

##Recode education variable Q35, as college or no college


dataset <- dataset %>%
  mutate(college = case_when(
    Q35 %in% c("Associate degree in college (2-years)", "Bachelor's degree in college (4-years)", 
               "Doctoral degree", "Master's degree", "Professional degree (JD, MD)") ~ "college",
    Q35 %in% c("High school graduate (high school diploma or equivalent including GED)", "Less than high school degree", "Some college but no degree") ~ "nocollege",
    TRUE ~ NA_character_
  ))

levels(dataset$college)=c("college", "nocollege")
describe(dataset$college)

  #save to design dataset, if we needed weights we would put them here.
datasetD <- svydesign(id=~1, data=dataset)
```


In this section I run a logistic regression on my newly transformed data. 

Unfortunately due to the small sample size of the survey I was provided in class to do this regression, the results are statistically insignificant, however, I will still analyze as though they are not. 
```{r}
#logit model
trump.logit <- svyglm(trumpvote~broke+partyid+college, design=datasetD, family="quasibinomial")
summary(trump.logit)
```


Below I will graph the predicted probabilities of each of the 8 possible respondents (binary DV and IVs mean 8 possible choices). Then I make a pie chart of my sample to give some insight on the demographics.
```{r}
##prediction frames

#We recoded the two IVs into two binary variables, this prediction frame tests all 8 possible outcomes and the probability a respondent would vote for Trump given the IVs
levels(dataset$broke)
prediction.frame.1 = expand.grid(broke = levels(dataset$broke),
                                 partyid = "democrat",
                                 college = levels(dataset$college))
prediction.frame.1
prediction.frame.2 = expand.grid(broke = levels(dataset$broke),
                                 partyid = "notdemocrat",
                                 college = levels(dataset$college))
prediction.frame.2

#you would have two predictions & two prediction frame
?predict.glm
predictions.1 = predict.glm(trump.logit, newdata = prediction.frame.1,
                            type = "response", se.fit = T,
                            interval = "confidence", level = .95)
predictions.2 = predict.glm(trump.logit, newdata = prediction.frame.2,
                            type = "response", se.fit = T,
                            interval = "confidence", level = .95)

predictions.1$fit
predictions.2
predictions.1


#This plots the predicted probability of those prediction frames we made above


plot(1:4, y = predictions.1$fit, ylim = c(0,1), axes = F,
     xlab = "View of Institutions in the U.S. and Level of Education",
     ylab = "Probability of Trump Vote",
     main =
       "Figure 3:View of Insitituions and Education Level Predicting \n the Probability for a Trump Vote",
     pch =16,
     col = c("lightblue", "lightblue", "lightblue", "lightblue"))
points(1:4, y = predictions.2$fit, pch =16,
       col = c("palevioletred", "palevioletred",
               "palevioletred", "palevioletred"))
segments(x0=1:4 , y0=predictions.1$fit-1.96*predictions.1$se.fit ,
         x1=1:4 , y1=predictions.1$fit+1.96*predictions.1$se.fit)
segments(x0=1:4 , y0=predictions.2$fit-1.96*predictions.2$se.fit ,
         x1=1:4 , y1=predictions.2$fit+1.96*predictions.2$se.fit)
axis(side=1, at=1:4, labels=c("College Educated \n Institutions Broken", 
                              "College Educated \n Institutions Not Broken", 
                              "Not College Educated \n Institutions  Broken", 
                              "Not College Educated \n Institutions Not Broken"))
axis(side=2)
legend("topright", legend = c("Democrat", "Republican"),
      col = c("lightblue", "palevioletred"), pch = 16, title = "Party")







#pie charts, descriptive stats about or population
 pie_old <- data.frame(response = c("Neither agree nor disagree ", "Somewhat agree", 
                                        "Strongly agree", "Somewhat disagree", "Strongly disagree"),
                          proportion = c(.196, 0.349, .387, .048, .02))

  pie_chart_old <- ggplot(pie_old, aes(x = "", y = proportion, fill = response)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar(theta = "y") +
    theme_minimal() +
    labs(title = "Figure1: Orginial proprtions of view of institutions in the U.S.") +
    geom_text(aes(label = paste0(round(proportion * 100, 1), "%\n(", round(proportion, 3), ")")),
              position = position_stack(vjust = 0.5)) +
    theme(legend.position = "bottom")  # Optional: Move legend to the bottom
  pie_chart_old
  
  #new pie chart
  
  pie_new <- data.frame(response = c("Agree", "Disagree"),
                     proportion = c(0.735, 0.265))
  
  # Create a pie chart
  pie_chart_new <- ggplot(pie_new, aes(x = "", y = proportion, fill = response)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar(theta = "y") +
    theme_minimal() +
    labs(title = "Figure 2:Transformed proprtions of view of insitutions in the U.S.") +
    geom_text(aes(label = paste0(round(proportion * 100, 1), "%\n(", round(proportion, 3), ")")),
              position = position_stack(vjust = 0.5)) +
    theme(legend.position = "bottom")  # Optional: Move legend to the bottom

  


```
